---
title: "Data 607 Final Project"
author: "Natalie Kalukeerthie, Anna Moy, Bishoy Sokkar"

output: 
  prettydoc::html_pretty:
    theme: architect
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

During the year of 2020, the world went into lock down, many people were forced to stay home, thus having the time to consume more media than ever before.

We could see movies being a popular source of entertainment during both the peak of lock down (Spring/Summer 2020) where people watched many movies in their homes and once restrictions were eased and people were allowed to finally congregate in public spaces again.

As the movie and entertainment industry have a large impact on today’s pop culture, we’re intrigued to see how successful certain movies were over the years, especially in 2019 and 2020. We will also be delving into the gross income of a variety of movies in order to find which movie was the most profitable. Comparing the 2019 and 2020 data to determine if the pandemic impacted the total profits theses movies received. 

Our datasets will be coming from Kaggle: [Movie Industry](https://www.kaggle.com/datasets/danielgrijalvas/movies) and through web scraping on [Box Office Mojo](https://www.boxofficemojo.com/year/2020/). 

Our team used slack, discord and github in order to collaborate effectively.

# Data Collection
``` {r load library, message = FALSE}
#load library
library(tidyverse)
library(rvest)
library(lubridate)
library(sf)
library(ggrepel)
library(countrycode)
library(rnaturalearth)
```

```{r read-csv-file, message = FALSE}
#read csv file from kaggle
movie <- read_csv("https://raw.githubusercontent.com/AnnaMoy/Data-607-Project/main/movies.csv")
movie
```

```{r scrape-movie}
# Webscraping from Boxoffice Mojo website
# link to the website
url <- "https://www.boxofficemojo.com/year/2020/"

# read url link 
webpage <- read_html(url)

# pull in movie title, gross, total gross and release date from 2020 movie information 
gross_2020 <- html_nodes(webpage, ".mojo-field-type-date , .mojo-estimatable , .mojo-cell-wide") %>%
  html_text()

# put the information into a dataframe
df <- data.frame(gross = gross_2020)

head(df,20)
``` 

We decided to get our dataset through web-scraping. By using the function html_nodes(), we pulled in gross movie data for 2020 from https://www.boxofficemojo.com/year/2020/ 

# Data Transformation and Cleaning

The originally scraped data is not in a usable format, so we must transform it into wide data so each column has it's own observations in each column. 

```{r data-transformation-webscrape}
# All the data was in one column therefore we split up the data into their own columns and provided new column names
df <- df %>%
  mutate(ind = rep(c("title", "gross_data", "total_gross","date"),length.out = n())) %>%
  group_by(ind) %>%
  mutate(id = row_number()) %>%
  spread(ind, gross) %>%
  select(-id)

#remove line 1 which initially has the old titles
moviedf <- df[-1,]


```

```{r data-transformation}
# Transform the release date in the kaggle movie dataset
movie<- movie %>% 
  separate_wider_delim(released, delim =" (", names= c("date","location"))

#convert release date into date format
movie$date <- mdy(movie$date)

#add in year and quarter
movie<-movie %>%
  mutate(qtr= quarter(date,with_year = T))
```

```{r join-scrap_data-kaggle_data}
#joining the data for kaggle movie and webscraped data from Box Office Mojo
movie_df <-inner_join(movie, moviedf, by=c("name" = "title"))

```

```{r making 2020 gross numeric}
#creating a subset and filter for 2020 movies
movie2020 <- movie_df %>%
  filter(year %in% c("2020"))

movie2020_numeric <- movie2020

#converting gross income into a numeric value for analysis
movie2020_numeric$total_gross_numeric <- movie2020_numeric$total_gross
movie2020_numeric$total_gross_numeric <- as.character(gsub("\\$", "", movie2020_numeric$total_gross_numeric))
movie2020_numeric$total_gross_numeric <- as.numeric(gsub(",", "", movie2020_numeric$total_gross_numeric))

#displaying highest gross incomes first
movie2020_numeric <- movie2020_numeric[order(movie2020_numeric$total_gross_numeric, decreasing = TRUE),]
movie2020_numeric

```

```{r making 2019 gross numeric2}
#creating a subset and filter for 2019 movies
movie2019 <- movie_df %>%
  filter(year %in% c("2019"))

movie2019_numeric <- movie2019

#converting gross income into a numeric value for analysis
movie2019_numeric$total_gross_numeric <- movie2019_numeric$total_gross
movie2019_numeric$total_gross_numeric <- as.character(gsub("\\$", "", movie2019_numeric$total_gross_numeric))
movie2019_numeric$total_gross_numeric <- as.numeric(gsub(",", "", movie2019_numeric$total_gross_numeric))

#displaying highest gross incomes first
movie2019_numeric <- movie2019_numeric[order(movie2019_numeric$total_gross_numeric, decreasing = TRUE),]
movie2019_numeric
```
# Analysis

Using our tidied data, we will use it to discover some insights on the movies released in 2020, such as:

Which movie in 2019 and 2020 was the most profitable?

To define a movie's success, I will measure it by profit, where if a movie makes more than what they spend for their budget, they will be considered successful, and the movie with the largest profit will be the most successful.

```{r ggplot 2020 and 2019,figures-side, fig.show="hold", out.width="50%", echo = FALSE}
ggplot(movie2020_numeric, aes(x = budget, y = total_gross_numeric)) +
  geom_point() +
  labs(title = "Budget vs Total Gross for Movies (2020)",
       x = "Budget (in million $)",
       y = "Total Gross (in million $)") +
  theme_minimal()

ggplot(movie2019_numeric, aes(x = budget, y = total_gross_numeric)) +
  geom_point() +
  labs(title = "Budget vs Total Gross for Movies (2019)",
       x = "Budget (in million $)",
       y = "Total Gross (in million $)") +
  theme_minimal()
```

Using a scatterplot to compare budget against gross_income of each movie, we can that there isn't a general trend here, however, there are a few movies that had similar budgets but varied in their overall gross.

Now that we see the trend, let's see which movies made a profit:

```{r,figures-side, fig.show="hold", out.width="50%", echo = FALSE}
#finding difference
movie2020_numeric$profit <- movie2020_numeric$total_gross_numeric - movie2020_numeric$budget

# Create a new column indicating whether the movie made a profit or not
movie2020_numeric$profitable <- ifelse(movie2020_numeric$profit > 0, "Profitable", "Not Profitable")

# Create the visualization
ggplot(movie2020_numeric, aes(x = budget, y = total_gross_numeric, color = profitable)) +
  geom_point() +
  labs(title = "Budget vs Total Gross for Movies (2020)",
       x = "Budget (in million $)",
       y = "Total Gross (in million $)",
       color = "Profitability") +
  theme_minimal()

#finding difference
movie2019_numeric$profit <- movie2019_numeric$total_gross_numeric - movie2019_numeric$budget

# Create a new column indicating whether the movie made a profit or not
movie2019_numeric$profitable <- ifelse(movie2019_numeric$profit > 0, "Profitable", "Not Profitable")

# Create the visualization
ggplot(movie2019_numeric, aes(x = budget, y = total_gross_numeric, color = profitable)) +
  geom_point() +
  labs(title = "Budget vs Total Gross for Movies (2019)",
       x = "Budget (in million $)",
       y = "Total Gross (in million $)",
       color = "Profitability") +
  theme_minimal()
```

From this visual we can identify which movies had a total gross that surpassed their budget, thus making a profit. We can also see that a majority of movies did not make a profit.



```{r}
# Sort the dataset by profit in descending order
movie2020_numeric_sorted <- movie2020_numeric[order(-movie2020_numeric$profit), ]

# We'll look at the top 5 movies
top_5_profitable <- head(movie2020_numeric_sorted, 5)

print(top_5_profitable[, c("name", "budget", "total_gross", "profit")])
```
From listing the top 5 movies, we can see that Bad Boys for Life was the most successful movie in 2020.

We can also compare the profits between 2019 and 2020 in our analysis. 

```{r 2019 profitable}
#finding difference
movie2019_numeric$profit <- movie2019_numeric$total_gross_numeric - movie2019_numeric$budget

# Create a new column indicating whether the movie made a profit or not
movie2019_numeric$profitable <- ifelse(movie2019_numeric$profit > 0, "Profitable", "Not Profitable")


# Sort the dataset by profit in descending order
movie2019_numeric_sorted <- movie2019_numeric[order(-movie2019_numeric$profit), ]

# We'll look at the top 5 movies
top_5_profitable_2019 <- head(movie2019_numeric_sorted, 5)

print(top_5_profitable_2019[, c("name", "budget", "total_gross", "profit")])

```

From listing the top 5 movies, we can see that Frozen II was the most successful movie in 2019.

# Data Visualization Feature To Create Map

Our team wanted to visualize the consumption of movies across the world, using ggplot along with the packages loaded below to create a map showing total movie revenue by country. To create a map we used dplyr to filter the relevant columns, including the country. We then grouped by country summing the revenue from each movie. We then used the Iso package to find a common code for each country to be easily identifiable. We then used the sf package to load a map and plot the data on. 

```{r}
map_data <- movie %>%
  select(year, country, budget, gross) %>%
  mutate( profit = gross-budget) %>%
  na.omit(gross) %>%
  filter(year <= 2010) %>%
  group_by(country) %>%
  summarise(total_income = sum(profit))


#Add ISO which is a 3 letters code for each country  for Join function to work
data_with_iso <- map_data %>%
  mutate(Iso3 = countrycode::countrycode(
    sourcevar = country,
    origin = "country.name",
    destination = "iso3c"))
#
world <- ne_countries(scale = "small", returnclass = "sf") 

countries_movie_consumption <- world %>%
  select(geometry, name, iso_a3) %>%
  left_join(data_with_iso, by = c("iso_a3" = "Iso3"))


world %>% 
  st_transform(crs = "+proj=robin") %>%
  ggplot() +
  geom_sf(color = "darkgray") +
  geom_sf(data = countries_movie_consumption, aes(fill = total_income)) +
  theme_minimal() +
   labs(
    title = "World Movie Revenue by Country from 2010 - 2020",
    subtitle = "Total Gross Income from Movies in million",
    fill = "Total Gross Income (USD)"
  )


```

From the plot above we can see that the USA is a lead in movie revenue. The USA movie industry not only generates substantial revenue but also serves as a key ambassador of American culture and values worldwide.


```{r}

map_data_2020 <- movie %>%
  select(year, country, budget, gross) %>%
  mutate( profit = gross-budget) %>%
  na.omit(gross) %>%
  filter(year == 2020) %>%
  group_by(country) %>%
  summarise(total_gross_income = sum(gross))

data_with_iso <- map_data_2020 %>%
  mutate(Iso3 = countrycode::countrycode(
    sourcevar = country,
    origin = "country.name",
    destination = "iso3c"))


world_2020 <- ne_countries(scale = "small", returnclass = "sf") 

countries_movie_consumption_2020 <- world_2020 %>%
  select(geometry, name, iso_a3) %>%
  left_join(data_with_iso, by = c("iso_a3" = "Iso3"))


world_2020 %>% 
  st_transform(crs = "+proj=robin") %>%
  ggplot() +
  geom_sf(color = "darkgray") +
  geom_sf(data = countries_movie_consumption_2020, aes(fill = total_gross_income)) +
  theme_minimal() +
   labs(
    title = "World Movie Revenue by Country from 2020",
    subtitle = "Total Gross Income from Movies in million",
    fill = "Total Gross Income (USD)"
  )

```

We can see from the plot above that in 2020 the movie revenue for three countries: USA, Canada and China. 

# Statistical Analysis
Null Hypothesis: There is no difference between 2019 and 2020 profit

Alternative Hypothesis: There is a difference between 2019 and 2020 profit

Reject the Null Hypothesis since p-value is less than .05.

```{r statistical-analysis}
#creating a subset and filter for 2019 and 2020 movies
movieall <- movie_df %>%
  filter(year %in% c("2019", "2020"))


#converting gross income into a numeric value for analysis
movieall$total_gross_numeric <- movieall$total_gross
movieall$total_gross_numeric <- as.character(gsub("\\$", "", movieall$total_gross_numeric))
movieall$total_gross_numeric <- as.numeric(gsub(",", "", movieall$total_gross_numeric))

#displaying highest gross incomes first
movieall2 <- movieall[order(movieall$total_gross_numeric, decreasing = TRUE),]


# 2019 and 2020 movie gross data
movieall3 <- movieall2 %>%
  mutate(profit = (total_gross_numeric- budget)) 

#T Test 
t.test(profit ~ year, data = movieall3)

#Make a copy of the year column
movieall3$yearchar = movieall3$year 
# change yearchar to categorical 
movieall3$yearchar <- as.factor(movieall3$yearchar)


#boxplot for year and profits
ggplot(movieall3, aes(yearchar, profit)) +
  geom_boxplot() +
  labs(
    title = "Movie Profit for 2019 and 2020",
    x = "Years",
    y = "Profit")
```


# Conclusion

We can see through our analysis that movies had a large impact on views, especially during the pandemic, where it became one of the main source of entertainment. The profits in movies in 2020 were significantly lower than those in 2019 due to the pandemic a lot less people went to the movies. 

We did come across a few challenges in the making of this project, one of them being choosing two datasets that have a relationship and can be used in analysis. It took us quite some time to find two datasets, especially since the two datasets could not be from the same source. Despite this difficulty, taking the time to discover and compare multiple datasets helped us think and interpret data in a relational sense.


